Road User video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda

### 2024-07-29 paper.qmd
Minor updates

---
title: "Road User video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
bibliography: [references.bib, opsnap.bib]
editor:
  markdown: 
    wrap: sentence
author:
  - name: Graham Farrell
    corresponding: true
    affiliations:
      - name: University of Leeds
        department: School of Law
        address: The Liberty Building
        city: Leeds
        country: United Kingdom
        postal-code: LS2 9JT
    orcid: 0000-0002-3987-8457
  - name: Robin Lovelace
    affiliations:
      - name: University of Leeds
        department: Institute for Transport Studies
        address: University Road
        city: Leeds
        country: United Kingdom
        postal-code: LS2 9JT
      - name: Active Travel England
        # department: Data and Digital
        # address: West Offices Station Rise
        # city: York
        # country: United Kingdom
        # postal-code: YO1 6GA
    orcid: 0000-0001-5679-6536
  - name: Steve O'Hern
    affiliations:
      - name: University of Leeds
        department: Institute for Transport Studies
        address: University Road
        city: Leeds
        country: United Kingdom
        postal-code: LS2 9JT
      - name: Monash University Accident Research Centre
        address: Clayton, VIC 3800, Australia
    orcid: 0000-0002-7961-3875
keywords: 
  - Road safety
  - Video evidence
  - Dashcam
  - Near misses
date: last-modified
format:
  # html: default
  # docx: default
  # pdf: default
  arxiv-pdf:
    keep-tex: true
linenumbers: true # Add (continuous) line numbers?
doublespacing: false # Double space the PDF output?
runninghead: "Preprint" # The text on the top of each page of the output
authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
editor_options: 
  markdown: 
    wrap: sentence
execute: 
  echo: false
  warning: false
  cache: false
---

@@

```{r}
#| eval: false
#| echo: false
quarto::quarto_render("paper.qmd", output_format = "pdf")
file.rename("paper.pdf", "paper-v10.pdf")
quarto::quarto_render("paper.qmd", output_format = "docx")
# file.rename("paper.docx", "/OneDrive/opsnap_leeds/paper-v10.docx")
file.rename("paper.docx", "paper-v10.docx")
system("gh release upload v1 paper-v10.docx paper-v10.pdf")
browseURL("paper.docx")
# Install extension:
system("quarto install extension mikemahoney218/quarto-arxiv")
system("sudo apt install lmodern")
```

@@

There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations shown in @tbl-locations.
The locations were scrambled by West Yorkshire Police before being made available for data protection purposes, which meant spatial analysis would only possible for ‘all’ video submissions, ignoring different types of road users or offence types.

```{r}
#| echo: false
#| label: tbl-locations
#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
# d_with_location |>
#   count(location, sort = TRUE) |>
#   mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
#   arrange(desc(n)) |>
#   head(10) |>
#   rename_all(snakecase::to_title_case) |> 
#   knitr::kable()
tbl = d_with_location |>
  count(location, sort = TRUE) |>
  mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
  arrange(desc(n)) |>
  head(10) |>
  rename_all(snakecase::to_title_case) |>
  rename(`Number` = N, `Percent` = `Percent of Records`) |>
  gt() |>
  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
  tab_style(
    style = cell_text(size=px(18)),
    locations = cells_body(columns = Location)
  ) |>
  gt_theme_espn()
# gt::gtsave(tbl, "tbl-locations.png")
knitr::include_graphics("tbl-locations.png")
```

@@

There was preliminary confirmation of this pattern in our analysis which found that offence types reported by vehicle drivers are systematically different to those of other road users.
Hence future research should consider that investigative and preventive approaches should be tailored to different contexts and types of road user.
Rural roads account for a disproportionate amount of fatalities @brakeDirectLineBrake2018, and further research using video evidence may inform preventive approaches tailored to road type.
Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing (see e.g., (@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010).

@@

>>>>

---
title: "Road User video evidence of road traffic offences: Preliminary analysis of Operation Snap data and suggestions for a research agenda"
bibliography: [references.bib, opsnap.bib]
editor:
  markdown: 
    wrap: sentence
keywords: 
  - Road safety
  - Video evidence
  - Near misses
  - Operation Snap
  - Dangerous driving
  - Antisocial driving
date: last-modified
format:
  # html: default
  # docx: default
  # pdf: default
  arxiv-pdf:
    keep-tex: true
linenumbers: true # Add (continuous) line numbers?
doublespacing: false # Double space the PDF output?
runninghead: "Preprint" # The text on the top of each page of the output
authorcols: true # Should authors be listed in a single column (default) or in multiple columns (`authorcols: true`)
editor_options: 
  markdown: 
    wrap: sentence
execute: 
  echo: false
  warning: false
  cache: false
---

@@

```{r}
#| eval: false
#| echo: false
# quarto::quarto_render("paper.qmd", output_format = "pdf")
# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-no-line-numbers-v11.pdf")
# quarto::quarto_render("paper.qmd", output_format = "arxiv-pdf")
# file.copy("paper.pdf", "paper-v11.pdf")
# file.copy("paper.pdf", "~/OneDrive/opsnap_leeds/paper-v11.pdf")
# quarto::quarto_render("paper.qmd", output_format = "docx")
# # file.copy("paper.docx", "~/OneDrive/opsnap_leeds/paper-v11.docx")
# file.copy("paper.docx", "paper-v11.docx")
# quarto::quarto_render("title.qmd", output_format = "arxiv-pdf")
# browseURL("title.pdf")
# system("gh release upload v1 --clobber paper-v11.pdf title.pdf")
# browseURL("paper.docx")
# # Install extension:
# system("quarto install extension mikemahoney218/quarto-arxiv")
# system("sudo apt install lmodern")
```

@@

There were `r unique(d_with_location$location) |> length()` unique locations (addresses) in the data, with the most common locations corresponding to busy roads: Meanwood Road (Leeds), Dewsbury Road (Wakefield) and Chapeltown Road (Leeds), with no single address accounting for more than 0.4% of records.
The locations were scrambled by West Yorkshire Police before being made available for data protection purposes, which meant spatial analysis would only possible for ‘all’ video submissions, ignoring different types of road users or offence types.

```{r}
#| echo: false
#| eval: false
#| label: tbl-locations
#| tbl-cap: "Most common locations recorded in the OpSnap dataset"
# d_with_location |>
#   count(location, sort = TRUE) |>
#   mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
#   arrange(desc(n)) |>
#   head(10) |>
#   rename_all(snakecase::to_title_case) |> 
#   knitr::kable()
tbl = d_with_location |>
  count(location, sort = TRUE) |>
  mutate(percent_of_records = round(n / nrow(d_with_location), 3) * 100) |>
  arrange(desc(n)) |>
  head(10) |>
  rename_all(snakecase::to_title_case) |>
  rename(`Number` = N, `Percent` = `Percent of Records`) |>
  gt() |>
  gt_plt_bar_pct(column = `Percent`, fill = "#252525", scaled = TRUE, labels=TRUE, width=120, height=30, font_size = "15px") |> 
  cols_width(1 ~ px(400), 2 ~ px(80)) |> 
  tab_style(
    style = cell_text(size=px(18)),
    locations = cells_body(columns = Location)
  ) |>
  gt_theme_espn()
# gt::gtsave(tbl, "tbl-locations.png")
# knitr::include_graphics("tbl-locations.png")
```

@@

There was preliminary confirmation of this pattern in our analysis which found that offence types reported by vehicle drivers are systematically different to those of other road users.
Hence future research should consider that investigative and preventive approaches should be tailored to different contexts and types of road user.
Rural roads account for a disproportionate amount of fatalities @brakeDirectLineBrake2018, and further research using video evidence may inform preventive approaches tailored to road type.
Policing and prevention efforts focused on where problems are concentrated are more resource efficient and is the foundation of problem-oriented policing [@laingAggressiveDriving2010; @scottSpeedingResidentialAreas2010].

@@

# Acknowledgements

[To be added after peer review]

<!--

We would like to thank Ian King for making the connections that led to this paper being written.
Thanks to Roger Beecham who wrote code for visualising some of the tables/figures in the paper, testing and supporting with reproducibility. 
Thanks to Simon D'vali and colleagues from City of Bradford Metropolitan Council for their support and advice on the project and colleagues from the Safer Cycling Yorkshire group.

-->
